### 2018-03-19 icfp18/supplement/icfp18-supplement.tex
added supplement

\PassOptionsToPackage{svgnames,dvipsnames,svgnames}{xcolor}

\documentclass[12pt]{report}
\newif\ificfp
\icfptrue
\newcommand{\todolater}[1]{{\color{magenta} TODO (Later): #1}}
\newcommand{\todo}[1]{{\color{red} TODO: #1}}

% some useful packages
\usepackage{times}     % use times font for document
%\usepackage{lmodern}
\usepackage{newtxtt}
\usepackage{bbm}
%\renewcommand{\ttdefault}{txtt} % use txtt for typewriter font
\usepackage{mathpazo}
\usepackage{mathpartir} % use package for inference rules
\usepackage{upgreek} % package for alternative greek letters (\uppi)
\usepackage{fullpage}
\usepackage{colortab}
%\usepackage{graphicx}
\usepackage[labelfont=bf]{caption}

% \usepackage{cleveref}
%\usepackage{MnSymbol}
\usepackage{fancyvrb}
\usepackage{microtype}
\usepackage[framemethod=tikz]{mdframed}

% Get just llangle and rrangle from MnSymbol
\makeatletter
\DeclareFontFamily{OMX}{MnSymbolE}{}
\DeclareSymbolFont{MnLargeSymbols}{OMX}{MnSymbolE}{m}{n}
\SetSymbolFont{MnLargeSymbols}{bold}{OMX}{MnSymbolE}{b}{n}
\DeclareFontShape{OMX}{MnSymbolE}{m}{n}{
    <-6>  MnSymbolE5
   <6-7>  MnSymbolE6
   <7-8>  MnSymbolE7
   <8-9>  MnSymbolE8
   <9-10> MnSymbolE9
  <10-12> MnSymbolE10
  <12->   MnSymbolE12
}{}
\DeclareFontShape{OMX}{MnSymbolE}{b}{n}{
    <-6>  MnSymbolE-Bold5
   <6-7>  MnSymbolE-Bold6
   <7-8>  MnSymbolE-Bold7
   <8-9>  MnSymbolE-Bold8
   <9-10> MnSymbolE-Bold9
  <10-12> MnSymbolE-Bold10
  <12->   MnSymbolE-Bold12
}{}

\let\llangle\@undefined
\let\rrangle\@undefined
\DeclareMathDelimiter{\llangle}{\mathopen}%
                     {MnLargeSymbols}{'164}{MnLargeSymbols}{'164}
\DeclareMathDelimiter{\rrangle}{\mathclose}%
                     {MnLargeSymbols}{'171}{MnLargeSymbols}{'171}
\makeatother

% \usepackage{rotating}
% \usepackage{pdflscape}

\usepackage[colorlinks=true,allcolors=Blue,backref,pageanchor=true,plainpages=false, pdfpagelabels, bookmarks,bookmarksnumbered,
pdfborder={0 0 0},  %removes outlines around hyper links in online display
]{hyperref}

\usepackage{amsmath,amssymb, amsthm}

\allowdisplaybreaks[1]
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{condition}[theorem]{Condition}

\usepackage{pfsteps}

\usepackage[noabbrev]{cleveref}

\newenvironment{proof-sketch}{\noindent{\emph{Proof Sketch.}}}{\qed}

% \makeatletter
% \renewenvironment{proof}[1][\proofname]{\par
%   \vspace{-\topsep}% remove the space after the theorem
%   \pushQED{\qed}%
%   \normalfont
%   \topsep0pt \partopsep0pt % no space before
%   \trivlist
%   \item[\hskip\labelsep
%         \itshape
%     #1\@addpunct{.}]\ignorespaces
% }{%
%   \popQED\endtrivlist\@endpefalse
%   \addvspace{6pt plus 6pt} % some space after
% }
% \makeatother
% \makeatletter
% \renewenvironment{proof-sketch}[1][\proofname]{\par
%   \vspace{-\topsep}% remove the space after the theorem
%   \pushQED{\qed}%
%   \normalfont
%   \topsep0pt \partopsep0pt % no space before
%   \trivlist
%   \item[\hskip\labelsep
%         \itshape
%     Proof Sketch\@addpunct{.}]\ignorespaces
% }{%
%   \popQED\endtrivlist\@endpefalse
%   \addvspace{6pt plus 6pt} % some space after
% }
% \makeatother
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions

\usepackage{mathtools}
\usepackage{ stmaryrd }
\usepackage[numbers,sort]{natbib}

% \usepackage{subfigure}

% Approximately 1" margins, more space on binding side
%\usepackage[letterpaper,twoside,vscale=.8,hscale=.75,nomarginpar]{geometry}
%for general printing (not binding)
\usepackage[letterpaper,twoside,vscale=.8,hscale=.75,nomarginpar,hmarginratio=1:1]{geometry}

%% Listings
\definecolor{mygray}{rgb}{0.75,0.75,0.75}
\usepackage{listings}
\lstset{tabsize=2, 
basicstyle=\ttfamily\fontsize{8pt}{1em}\selectfont, 
commentstyle=\itshape\ttfamily\color{gray},
stringstyle=\ttfamily\color{purple},
mathescape=false,escapeinside={(~}{~)},
numbers=left, numberstyle=\scriptsize\color{mygray}, language=ML,showspaces=false,showstringspaces=false,xleftmargin=0pt, numbersep=-6pt, morekeywords=[1]{try,spliced,tyfam,opfam,let,fn,val,def,casetype,objtype,metadata,of,*,string,lexer,parser,datatype,new,toast,notation,module,switch,where,expansions,require,import,for,ana,syn,opcon,tycon,metasignature,metamodule,metasig,metamod,static,at,by,tycase,mod,macro,match,pattern,in,patterns,expressions,implicit,forall,int,exptsm,pattsm},deletekeywords={double,structure,of},classoffset=0, xleftmargin=0pt, 
aboveskip=3pt,belowskip=2pt,
moredelim=**[is][\color{red}]{SSTR}{ESTR},
moredelim=**[is][\color{Periwinkle}]{SHTML}{EHTML},
moredelim=**[is][\color{purple}]{SCSS}{ECSS},
moredelim=**[is][\color{brown}]{SSQL}{ESQL},
moredelim=**[is][\color{orange}]{SCOLOR}{ECOLOR},
moredelim=**[is][\color{magenta}]{SPCT}{EPCT}, 
moredelim=**[is][\color{gray}]{SNAT}{ENAT}, 
moredelim=**[is][\color{Periwinkle}]{SURL}{EURL},
moredelim=**[is][\color{Sealavender}]{SQT}{EQT},
moredelim=**[is][\color{Periwinkle}]{SGRM}{EGRM},
moredelim=**[is][\color{Yellowlavender}]{SID}{EID},
moredelim=**[is][\color{Sepia}]{SUS}{EUS},
moredelim=**[is][\color{black}]{SOK}{EOK},
% deletestring=[d]{"},
}
\lstset{morecomment=[n]{/*}{*/}}

\newcommand{\liv}[1]{\lstinline{#1}}
\newcommand{\li}[1]{\lstinline[basicstyle=\ttfamily\fontsize{10pt}{1em}\selectfont]{#1}}
\newcommand{\lismall}[1]{\lstinline[basicstyle=\ttfamily\fontsize{9pt}{1em}\selectfont]{#1}}
\newcommand{\lifootnote}[1]{\lstinline[basicstyle=\ttfamily\fontsize{7pt}{1em}\selectfont]{#1}}

% http://tex.stackexchange.com/q/43526
% fix the apparently deliberate but undocumented behaviour of disabling escapes other than mathescape in TextStyle (used by \lstinline)
% there may be a good reason why this is disabled by default, so beware in case it causes any problems
\usepackage{etoolbox}
\makeatletter
\patchcmd{\lsthk@TextStyle}{\let\lst@DefEsc\@empty}{}{}{\errmessage{failed to patch}}
\makeatother

% Provides a draft mark at the bottom of the document. 
% \draftstamp{\today}{DRAFT}

% I hate hyphenation.
%\lefthyphenmin=5
\definecolor{light-gray}{gray}{0.95}

\input{macros}

% allow interrupted equation numbering
% taken from http://tex.stackexchange.com/questions/101002/interrupting-and-resuming-subequations
% \makeatletter
% \def\user@resume{resume}
% \def\user@intermezzo{intermezzo}
% %
% \newcounter{previousequation}
% \newcounter{lastsubequation}
% \newcounter{savedparentequation}
% \setcounter{savedparentequation}{1}
% % 
% \renewenvironment{subequations}[1][]{%
%       \def\user@decides{#1}%
%       \setcounter{previousequation}{\value{equation}}%
%       \ifx\user@decides\user@resume 
%            \setcounter{equation}{\value{savedparentequation}}%
%       \else  
%       \ifx\user@decides\user@intermezzo
%            \refstepcounter{equation}%
%       \else
%            \setcounter{lastsubequation}{0}%
%            \refstepcounter{equation}%
%       \fi\fi
%       \protected@edef\theHparentequation{%
%           \@ifundefined {theHequation}\theequation \theHequation}%
%       \protected@edef\theparentequation{\theequation}%
%       \setcounter{parentequation}{\value{equation}}%
%       \ifx\user@decides\user@resume 
%            \setcounter{equation}{\value{lastsubequation}}%
%          \else
%            \setcounter{equation}{0}%
%       \fi
%       \def\theequation  {\theparentequation  \alph{equation}}%
%       \def\theHequation {\theHparentequation \alph{equation}}%
%       \ignorespaces
% }{%
% %  \arabic{equation};\arabic{savedparentequation};\arabic{lastsubequation}
%   \ifx\user@decides\user@resume
%        \setcounter{lastsubequation}{\value{equation}}%
%        \setcounter{equation}{\value{previousequation}}%
%   \else
%   \ifx\user@decides\user@intermezzo
%        \setcounter{equation}{\value{parentequation}}%
%   \else
%        \setcounter{lastsubequation}{\value{equation}}%
%        \setcounter{savedparentequation}{\value{parentequation}}%
%        \setcounter{equation}{\value{parentequation}}%
%   \fi\fi
% %  \arabic{equation};\arabic{savedparentequation};\arabic{lastsubequation}
%   \ignorespacesafterend
% }
% \makeatother

\begin{document} 
\title{Reasonably Programmable Literal Notation\\ (ICFP 2018 Anonymous Supplemental Material)}
\date{March 16, 2018}
\maketitle
\tableofcontents
\appendix

\chapter{Additional Examples}

\begin{figure*}[t]
\begin{subfigure}[t]{0.50\textwidth}
\begin{lstlisting}
  module Html = {
    /* a simplified encoding of HTML */
    type tag = string;
    type attr = (string, string);
    type attrs = list(attr);
    type t = 
      | Text(string) 
      | Elem(tag, attrs, list(t));
  }
\end{lstlisting}
\vspace{-4px}
\caption{The \li{Html} module, which defines \li{Html.t}.}
\label{fig:html-type-def}
\end{subfigure}
\hfill
\begin{subfigure}[t]{0.49\textwidth}
\begin{lstlisting}[morekeywords={expression}]
  module HtmlNotation {
    notation $html at Html.t {
      lexer Relit.TrivLexer;
      expression parser HtmlParser.expr;
      pattern parser HtmlParser.pat;
      expansions require 
        { module Html = Html; }
    }
  }
\end{lstlisting}
\vspace{-4px}
\caption{The \li{$html} expression and pattern TLMs.}
\label{fig:html-tlm-def}
\end{subfigure}

\begin{subfigure}[t]{\textwidth}
\vspace{5px}
% let pElem : Html.tag -> Html.attrs -> list(ProtoExpr.t) -> ProtoExpr.t = (* ... *)
% let pTextElem : Html.tag -> Html.attrs, string) -> ProtoExpr.t = (* ... *)
\begin{lstlisting}
  module HtmlParser = Relit.HandRolledExprPatParser({
    module ProtoHtml = {
      type t = PText(string)
             | PElem(Html.tag, Html.attrs, list(t))
             | PSplicedChildren(Html.tag, Html.attrs, Relit.Segment.t)
             | PSplicedElem(Relit.Segment.t) 
             | PSplicedText(Relit.Segment.t) 
      let parse : string => t = 
        /* ... via Markup.ml's stream combinators (see text) ... */;
    };
    module ExprExpander = {
      open notation Relit.$proto_expr; open ProtoHtml;
      let expand_attr = (x1, x2) => `(SURL ($s `(EURLx1SURL)`, $s `(EURLx2SURL)`)EURL )`;
      let expand_attrs = (attrs) => `(SURL $list `(EURLList.map(expand_attr, attrs)SURL)`EURL )`;
      let expand_proto = fun 
      | PText(text) => `(SURL Html.Text($s `(EURLtextSURL)`)EURL )` 
      | PElem(tag, attrs, children) => 
        `(SURL Html.Elem($s `(EURLtagSURL)`, `(EURLexpand_attrs(attrs)SURL)`, 
            $list `(EURLList.map(expand_parsed, children)SURL)`)EURL )`
      | PSplicedChildren(tag, attrs, seg) => 
        `(SURL Html.Elem($s `(EURLtagSURL)`, `(EURLexpand_attrs(attrs)SURL)`, 
            $spliced `(EURLsegSURL : list(Html.t))`)EURL )`
      | PSplicedHtml(seg) => `(SURL $spliced `(EURLsegSURL : Html.t) EURL)`);
      | PSplicedText(seg) => `(SURL Html.Text($spliced `(EURLsegSURL : striSURLEURLng)`) EURL)`
    };
    let expr = (body) => body |> ProtoHtml.parse |> ExprExpander.expand_proto;
    let pat  = (body) => expr(body) |> ProtoPat.from_expr;
  })
\end{lstlisting}
\vspace{-6px}
\caption{The implementation of \li{HtmlParser}, which defers the parsing step to the \li{Markup.ml} library \cite{markupml}.}
\label{fig:html-parser}
\end{subfigure}
\begin{subfigure}[t]{\textwidth}
\lstset{deletestring=[d]{"}}
\begin{lstlisting}
  open HtmlNotation; 
  exception UnexpectedFormat;
  let scrape_data = $html.(fun 
  | `(SURL<div... class="result">EURL
      [ `(SCOLOR<h1><$$>ECOLORtitleSCOLOR</$$></h1>ECOLOR)`, author, summary, ..._ ]SURL</div>EURL)` => 
    (title, author, summary)
  | _ => raise UnexpectedFormat);
\end{lstlisting}
\vspace{-4px}
\caption{Using a pattern TLM to deconstruct an HTML value.}
\label{fig:pattern-tlm-example}
\end{subfigure}
\vspace{2px}
\caption{Case Study: HTML literals as a library}
\label{fig:html-case-study}
\end{figure*}

\section{HTML Literals} 
Reason is seeing increasing adoption in the domain of client-side web programming, where generating and manipulating HTML data is central. 
There are many ways to represent HTML data in Reason, but one simple and useful representation is specified by \li{Html.t}, defined in Fig.~\ref{fig:html-type-def}. However, manually applying the constructors of this type is tedious and, for many web programmers, the induced notation is unfamiliar. It also makes it difficult to copy-and-paste from, for example, existing HTML files, as might be useful when refactoring an existing project to use Reason. Instead, we would like HTML literal notation. 
The \li{$html} TLM defined in Fig.~\ref{fig:html-tlm-def} implements the standard HTML notation extended with the tags \li{<$>} and \li{<$$>}, which support \li{Html.t} splicing and \li{string} splicing, respectively. It also supports the suffix \li{...} on a tag, which causes the body of the tag to be a splice of type \li{list(Html.t)}.\footnote{We would also in practice have splicing for  attribute values, but for simplicity we omit these.} The example in the paper demonstrated the first of these. The example in Fig.~\ref{fig:pattern-tlm-example} shows the list splicing form in pattern position.% HTML splicing form, \li{<$>}$e$\li{</$>}.

% The HTML standard is complex. In addition, good error handling is important in practice. As such, most practical HTML parsing libraries are ``hand-rolled'', i.e. they are not generated from a context-free grammar by a general parser generator like Menhir. For example, the robust \li{Markup.ml} library for OCaml provides a simple one-pass stream-based interface that can be used to generate an HTML representation of the user's choice \cite{ocaml}. Let us outline how we can use this library to implement HTML literals, with support for splicing as just specified. 
In implementing HTML literals, using a parser generator may not be ideal. Instead, ``post-processing'' the result of an existing parser is a good approach. Fig.~\ref{fig:html-parser} shows much of the implementation of \li{HtmlParser},  which is constructed by applying a functor that constructs a module that has the same essential interface as the modules that Menhir generates, with the functions \li{expr : string -> ProtoExpr.t} and \li{pat : string -> ProtoPat.t} of the input module exported as the ``non-terminals''. The companion lexer is \li{Relit.TrivLexer}. The implementation of \li{expr} on Line 23 of Fig.~\ref{fig:html-parser} reveals that the literal body is parsed by first calling \li{ProtoHtml.parse}, then passing the result on to \li{ExprExpander.expand_proto}.

The \li{ProtoHtml.parse} function, outlined on Lines 8-9, generates a value of the type \li{ProtoHtml.t} defined on Lines 2-6 and so named because it is similar to \li{Html.t} but distinguishes spliced segments, like a proto-expression. We omit the definition of \li{ProtoHtml.parse} but conceptually, it starts by creating a character stream from the literal body, then from that creates a \li{Markup.ml} HTML content stream, then transforms that to a stream that emits the signals from the HTML content stream until it sees a splice start tag (which is reported as a syntax error relative to the HTML standard), at which point it calls \li{Segment.read_to}, which was described in the paper, to advance the original character stream until it recognizes the closing tag, emitting the generated segment and then returning control to the HTML content stream. The final step is to fold over this stream to produce the desired result of type \li{ProtoHtml.t}.

The \li{ExprExpander.expand_proto} function defined, with some helper functions, on Lines~9-20 finishes the job of the TLM parser by mapping the parsed proto-HTML to a proto-expression using the \li{$proto_expr} TLM previously described (in this case, using the \li{$list} antiquotation form that produces a parse tree of OCaml's built in list literal form from a list of parse trees). It would be straightforward to reuse the \li{ProtoHtml} module for different variations on HTML encodings, implementing only the expander anew. Overall, the amount of code needed to implement HTML literals.

Reason currently builds in ``JSX'' literals, which also support an HTML-like notation with support for splicing in a somewhat idiosynchratic manner motivated by certain JavaScript libraries (in particular, React). The expansion of this notation must be interpretted by a PPX rewriter. The problem is that this makes it difficult to use different HTML-related libraries at once. For example, ReasonReact and an XML library might both want to use this notation. This approach solves this problem (as well as other problems with the PPX-based approach as discussed in the paper). % For example, another encoding is one where each tag in the HTML standard correspond to a distinct datatype constructor.

\section{Implementing Quasiquotation} The Reason quasiquotation TLMs, e.g. \li{$re_expr}, can similarly be implemented by the functional transformations outlined below, with an example from each step (grossly simplified from the actual \li{Parsetree} representation) on the right. In words, \li{$re_expr} first programmatically invokes the Reason parser on the literal body. It next serializes the generated parse tree to Reason source code, then parses that. This produces a parse tree that, if evaluated in the appropriate environment, will produce the original parse tree. The final step is to implement antiquotation as described above by repurposing the generalized literal forms in the body, using the source locations from the first parse tree, which have  been carried into the second parse tree as constants.
\begin{lstlisting}[numbers=none,basicstyle=\ttfamily\fontsize{8pt}{1em}\selectfont]
  body             "2 + `(xyz)`" 
  |> parse_re      Plus(Num(2, Loc(0)), GenLit("xyz", Loc(6)))
  |> serialize_re  "Plus(Num(2, Loc(0)), GenLit(\"SCSSxyz\ECSS", Loc(6)))"
  |> parse_re      Ap(V("Plus"), /*...*/Ap(V("GenLit"), Pair(Str("xyz"), /*...*/Num(6)))
  |> genlit_to_sp  Ap(V("Plus"), /*...*/Spliced(6,8,TyPath(["ProtoExpr","t"])))
\end{lstlisting}

% \section{Regular Expressions, More Abstractly}
% A regular expression, or \emph{regex}, is a description of a \emph{regular language} \cite{Thompson:1968:PTR:363347.363387}. Regexes arise with some frequency in fields like natural language processing and bioinformatics.

% \paragraph{Recursive Sums}
% One way to encode regular expressions in Reason is as values of the recursive datatype abbreviated \li{rx} in Figure \ref{fig:datatype-rx}.

% \begin{figure}[h]
% \begin{lstlisting}[numbers=none]
% type rx = Empty | Str(string) | Seq(rx, rx) 
%         | Or(rx, rx) | Star(rx);
% \end{lstlisting}
% \caption{Definition of the recursive labeled sum type \li{rx}}
% \label{fig:datatype-rx}
% \end{figure}
% We can construct a regex that matches the strings \li{"SSTRAESTR"}, \li{"SSTRTESTR"}, \li{"SSTRGESTR"} or \li{"SSTRCESTR"} (i.e. DNA bases) as follows:
% \begin{lstlisting}[numbers=none]
% Or(Str "SSTRAESTR", Or(Str "SSTRTESTR", Or(Str "SSTRGESTR", Str "SSTRCESTR")))
% \end{lstlisting}

% Given a value of type \li{rx}, we can deconstruct it by pattern matching. For example, the function \li{is_dna_rx} defined in Figure \ref{fig:is_dna_rx} detects regular expressions that match DNA sequences.

% \begin{figure}[h]
% \begin{lstlisting}[numbers=none]
% fun is_dna_rx(r : rx) : boolean => 
%   switch r { 
%   | Str "SSTRAESTR" => True;
%   | Str "SSTRTESTR" => True;
%   | Str "SSTRGESTR" => True;
%   | Str "SSTRCESTR" => True;
%   | Seq (r1, r2) => (is_dna_rx r1) andalso (is_dna_rx r2);
%   | Or  (r1, r2) => (is_dna_rx r1) andalso (is_dna_rx r2);
%   | Star(r') => is_dna_rx r';
%   | _ => False;
%   };
% \end{lstlisting}
% \caption{Pattern matching over regexes in Reason}
% \label{fig:is_dna_rx}
% \end{figure}

% \paragraph{Abstract Types} Encoding regexes as values of type \li{rx} is straightforward, but there are reasons why one might not wish to expose this encoding to clients directly.

% First, regexes are usually identified up to their reduction to a normal form. For example, \li{Seq(Empty, Str "SSTRAESTR")} has normal form \li{Str("SSTRAESTR")}. It can be useful for regexes with the same normal form to be  indistinguishable from the perspective of client code. (The details of regex normalization are not important for our purposes, so we omit them.)

% Second, it can be useful for performance reasons to maintain additional data alongside each regex (e.g. a corresponding finite automaton.) In fact, there may be many ways to represent regexes, each with different performance trade-offs, so we would like to provide a choice of representations behind a common interface.

% To achieve these goals, we turn to the OCaml module system, which is based closely on the SML module system \cite{mthm97-for-dart,dreyer2005understanding} (which originates in early work by MacQueen \cite{MacQueen:1984:MSM:800055.802036}.) In particular, let us define the {signature} abbreviated \li{RX} in Figure \ref{fig:signature-RX}.
% %Notice that it exposes an interface otherwise  to the one available using a case type:

% \begin{figure}[ht]
% \begin{lstlisting}[deletekeywords={case}]
% /* abstract regex unfoldings */
% type u('a) = UEmpty | UStr string | USeq ('a, 'a) 
%            | UOr ('a, 'a) | UStar of 'a;

% module type RX = {
%   type t (* abstract *);

%   (* constructors *)
%   let Empty : t;
%   let Str : string -> t;
%   let Seq : t * t -> t;
%   let Or : t * t -> t;
%   let Star : t -> t;

%   (* produces the normal unfolding *)
%   let unfold_norm : t -> u(t);
% end

% module R1 : RX = { (* ... *) };
% module R2 : RX = { (* ... *) };
% \end{lstlisting}
% \vspace{-5px}
% \caption{The \lstinline{RX} signature and two example implementations}
% \label{fig:signature-RX}
% \end{figure}

% The clients of any module \lstinline{R} that has been sealed by \lstinline{RX}, e.g. \li{R1} or \li{R2}  in Figure \ref{fig:signature-RX}, manipulate regexes as values of type \li{R.t} using the interface specified by \li{RX}. For example, a client can construct a regex matching DNA bases by projecting the value constructors out of \li{R} and applying them as follows:
% \begin{lstlisting}[numbers=none]
% R.Or(R.Str "SSTRAESTR", R.Or(R.Str "SSTRTESTR", R.Or (R.Str "SSTRGESTR", R.Str "SSTRCESTR")))
% \end{lstlisting}

% Because the identity of the representation type \lstinline{R.t} is held abstract by the signature, the only way for a client to construct a value of this type is through the values that \li{RX} specifies (i.e. we have defined an \emph{abstract data type (ADT)}  \cite{liskov1974programming}.) Consequently, representation invariants need only be established locally within each module.

% Similarly, clients cannot interrogate the structure of a value \li{r : R.t} directly. Instead, the signature specifies a function \li{R.unfold_norm} that produces the \emph{normal unfolding} of a given regex, i.e. a value of type \li{u(R.t)} that exposes only the outermost form of the regex in normal form (this normal form invariant is specified only as an unenforced side condition that implementations are expected to obey, as is common practice in languages like ML.) Clients can pattern match over the {normal unfolding} in the familiar manner, as shown in Figure \ref{fig:is_dna_rx_prime}. 
% \begin{figure}
% \begin{lstlisting}[numbers=none]
% fun is_dna_rx'(r : R.t) : boolean => 
%   switch R.unfold_norm r { 
%   | UStr "SSTRAESTR" => True;
%   | UStr "SSTRTESTR" => True;
%   | UStr "SSTRGESTR" => True;
%   | UStr "SSTRCESTR" => True;
%   | USeq (r1, r2) => (is_dna_rx' r1) && (is_dna_rx' r2);
%   | UOr (r1, r2) => (is_dna_rx' r1) && (is_dna_rx' r2);
%   | UStar r' => is_dna_rx' r';
%   | _ => False;
%   };
% \end{lstlisting}
% \vspace{-5px}
% \caption{Pattern matching over normal unfoldings of regexes}
% \label{fig:is_dna_rx_prime}
% \end{figure}

% The normal unfolding suffices in situations where a client needs to examine only the outermost structure of a regex. However, in general, a client may want to pattern match more deeply into a regex. There are various ways to approach this problem.

% One approach is to define auxiliary functions that construct $n$-deep unfoldings of \li{r}, where $n$ is the deepest level at which the client wishes to expose the normal structure of the regex. For example, it is easy to define a function \li{unfold_norm2 : R.t -> u(u(R.t))} in terms of \li{R.unfold_norm} that allows pattern matching to depth $2$.\footnote{Defining an unfolding \emph{generic} in $n$ is a more subtle problem that is beyond the scope of this work.}

% Another approach is to \emph{completely unfold} a value of type \li{t} by applying a function \li{view : R.t -> rx} that recursively applies \li{R.unfold_norm} to exhaustion. The type \li{rx} was defined in Figure \ref{fig:datatype-rx}.  Computing the complete unfolding (also called the \emph{view}) can have higher dynamic cost than computing an incomplete unfolding of appropriate depth, but it is also a simpler approach (i.e.   lower cognitive cost can justify higher dynamic cost.)

% \begin{figure}[t]
% \begin{lstlisting}[numbers=none]
% module RXUtil(R : RX) = {
%   let unfold_norm2(r : R.t) : u(u(R.t)) => /* ... */;

%   let view(r : R.t) : rx => 
%     switch R.unfold_norm r { 
%     | UEmpty => Empty;
%     | UStr s => Str s;
%     | USeq (r1, r2) => Seq (view r1, view r2);
%     | UOr (r1, r2) => Or (view r1, view r2);
%     | UStar r => Star (view r);
%     };

%   let parse_str(x : string) : R.t => /* ... */
% };
% \end{lstlisting}
% \vspace{-5px}
% \caption{The definition of \li{RXUtil}}
% \vspace{-5px}
% \label{fig:RXUtil}
% \end{figure}
% Typically, utility functions like \li{unfold_norm2} and \li{view} are defined in a \emph{functor} (i.e. a function at the level of modules) like \li{RXUtil} in Figure \ref{fig:RXUtil}, so that they need only be defined once, rather than separately for each module \li{R : RX}. The client can instantiate the functor by applying it to their choice of module as follows:
% \begin{lstlisting}[numbers=none]
% module RU = RXUtil(R)
% \end{lstlisting}

% \section{Using A Parser Generator to Define a TLM}
% Using these definitions, an implementation of a parser generator \li{P : PARSEGEN} and the static parametric TLMs \li{$grammar}, \li{$proto_typ} and \li{$proto_expr} as described in Sec. 6 of the paper, we might define a TLM for regexes (implementing a subset of the POSIX regex syntax for simplicity) as shown in Figure \ref{fig:rx-grammar-based}.

% The value \li{spliced_uexp} has type \li{proto_typ -> P.grammar(proto_expr)} and operates as described in Sec. 6 -- it recognizes the syntax of unexpanded expressions and produces the corresponding reference to a spliced expression, with the given type annotation.

% \begin{figure}[h!]
% \vspace{-5px}
% \begin{lstlisting}[deletekeywords={as}]
% (* we use regexes themselves to define lexemes in the grammar,
%  * so we need some static phase helpers. *)
% static module RS : RX = /* ... */;
% static module RU = RXUtil(RS);

% (* a TLM parametric over all modules R : RX *)
% syntax $rx(R : RX) at R.t by static {
%   P.generate ($grammar P proto_expr [|SHTML #\label{line:rx_parse_fn_start}#
%     start <- ""
%       EHTMLfun () => $proto_expr `SCSSR.EmptyECSS`SHTML
%     start <- "(" start ")"
%       EHTMLfun e => eSHTML
%     token str_tok #\label{line:str_tok_start}#
%       EHTML/* cannot use $rx within its own defn */
%       RU.parse_str "SSTR[^(@$]+ESTR" SHTML #\label{line:str_tok_end}#
%     start <- str_tok
%       EHTMLfun s => $proto_expr `SCSSR.Str %(ECSSstr_to_proto_str_lit sSCSS)ECSS`SHTML
%     start <- start start
%       EHTMLfun e1 e2 => $proto_expr `SCSSR.Seq (%ECSSe1SCSS, %ECSSe2SCSS)ECSS`SHTML
%     start <- start "|" start 
%       EHTMLfun e1 e2 => $proto_expr `SCSSR.Or (%ECSSe1SCSS, %ECSSe2SCSS)ECSS`SHTML
%     start <- start "*"
%       EHTMLfun e => $proto_expr `SCSSR.Star %ECSSe`SHTML
%     start <- "%{" ^(EHTMLspliced_uexp ($proto_typ `SCSSR.tECSS`)SHTML) "}" #\label{line:splicing-start}#
%       EHTMLfun e => eSHTML
%     start <- "${" ^(EHTMLspliced_uexp ($proto_typ `SCSSstringECSS`)SHTML) "}"
%       EHTMLfun e => $proto_expr `SCSSR.Str %(ECSSeSCSS)ECSS`SHTML #\label{line:splicing-end}#
%   EHTML|])
% }; #\label{line:rx_parse_fn_end}#
% \end{lstlisting}
% \vspace{-12px}
% \caption{A grammar-based definition of \texttt{\$rx}}
% \vspace{-15px}
% \label{fig:rx-grammar-based}
% \end{figure}

\input{appendix}

% \input{static-language}

\bibliographystyle{plainnat}
\bibliography{../../../../papers/research} %your bib file

\end{document}